### 2016-02-12 text/Chapter05.tex
Adding translation for implicits

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{num}): After applying $\ident{split}_{\cclrd{r}, \cclrd{s}\cseq\cclrd{t}}$, the types of 
  $\ctx_1, \ctx_2$ are $\ctyp{\cclrd{r}}(\ldots)$ and $\ctyp{\cclrd{s}\cseq\cclrd{t}}(\ldots)$, respectively.
  $g$ requires $\ctyp{\cclrd{s}}(\ldots)$ and so the result of $\ident{cobind}_{\cclrd{s},\cclrd{t}}$ is
  $\ctyp{\cclrd{t}}\tau_1$ as required by $f$.
\end{proof}

@@

In order to prove type safety, we first extend the canonical forms lemma
(Lemma~\ref{thm:semantics-fp-canon}) and the preservation under substitution lemma
(Lemma~\ref{thm:semantics-fp-pres-subst}). Those need to consider the new (\emph{df}) 
and (\emph{prev}) typing rules and substitution under the newly introduced expression forms 
$\ident{Df}\langle\ldots\rangle$ and $\ident{prev}_{\cclrd{n}}$. Then 
we extend the type preservation (Theorem~\ref{thm:semantics-fp-pres}) and progress
(Theorem~\ref{thm:semantics-fp-prog}).

@@

>>>>

\vspace{0.25em}\noindent\hangindent=0.6cm 
Case (\emph{app}): After applying $\ident{split}_{\cclrd{r}, \cclrd{s}\cseq\cclrd{t}}$, the types of 
  $\ctx_1, \ctx_2$ are $\ctyp{\cclrd{r}}(\ldots)$ and $\ctyp{\cclrd{s}\cseq\cclrd{t}}(\ldots)$, respectively.
  $g$ requires $\ctyp{\cclrd{s}}(\ldots)$ and so the result of $\ident{cobind}_{\cclrd{s},\cclrd{t}}$ is
  $\ctyp{\cclrd{t}}\tau_1$ as required by $f$.
\end{proof}

@@

In order to prove type safety, we first extend the canonical forms lemma
(Lemma~\ref{thm:semantics-fp-canon}) and the preservation under substitution lemma
(Lemma~\ref{thm:semantics-fp-pres-subst}). Those need to consider the new (\emph{df}) 
and (\emph{prev}) typing rules and substitution under the newly introduced expression forms 
$\ident{Df}\langle\ldots\rangle$ and $\ident{prev}_{\cclrd{n}}$. We show that the translation 
rule for \kvd{prev} produces well-typed expressions. Finally, we extend the type preservation 
(Theorem~\ref{thm:semantics-fp-pres}) and progress (Theorem~\ref{thm:semantics-fp-prog}).

\begin{theorem}[Well-typedness of the \kvd{prev} translation]
\label{thm:semantics-welltyped}
Given a typing derivation for a well-typed closed expression $\coctx{}{\cclrd{r}} \vdash e : \tau$,
the translated program $f$ obtained using the rules in Figure~\ref{fig:semantics-translation}
and Figure~\ref{fig:semantics-ext-df} is well-typed, \ie~in the target langauge: $\vdash f : \sem{\coctx{\Gamma}{\cclrd{r}}} \rightarrow \sem{\tau}$.
\end{theorem}
\begin{proof}
By rule induction over the derivation of the translation.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{var, num, abs, app}): As before.

\vspace{0.5em}\noindent\hangindent=0.6cm 
Case (\emph{prev}): Type of $\ctx$ is $\ctyp{\cclrd{n+1}}\tau$ and so we can apply the (\emph{prev}) rule.
\end{proof}

@@

\subsection{Coeffect langauge for implicit parameters}
\label{sec:semantics-proofs-impl}

::::TODO::::

\paragraph{Properties.}

\begin{figure}[h!]
\paragraph{Language syntax}

\begin{equation*}
\begin{array}{rcl}
v &=& \ldots \sep \ident{Impl}(v, \langle \ident{?p}_1\mapsto v_1, \ldots, \ident{?p}_n\mapsto v_n\rangle) \\
e &=& \ldots \sep \ident{Impl}(e, \langle\ident{?p}_1\mapsto e_1, \ldots, \ident{?p}_n\mapsto e_n\rangle) \\
  && \hspace{1.4em} \sep \ident{lookup}_{\cclrd{\ident{?p}}}~e \sep \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}~e_1~e_2\\
C &=& \ldots \sep \ident{lookup}_{\cclrd{\ident{?p}}}~\_ \sep \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}~\_~e \sep \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{r}}~v~\_
\end{array}  
\end{equation*}

\begin{equation*}
\tyrule{impl}
  { \Gamma \vdash e:\tau & \forall i\in\{ 1 \ldots n \}. ~ \Gamma \vdash e_i : \ident{num}}
  { \Gamma \vdash \ident{Impl}(e, \langle \ident{?p}_1\mapsto e_1, \ldots, \ident{?p}_n\mapsto e_n \rangle) : \ctyp{\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \}}}{\tau} }
\end{equation*}
\begin{equation*}
\tyrule{lookup}
  { \Gamma \vdash e : \ctyp{\cclrd{ \{ \ldots, \ident{?p}, \ldots \} }}{\tau} }
  { \Gamma \vdash \ident{lookup}_{\cclrd{\ident{?p}}}~e : \ident{num} }
\end{equation*}
\begin{equation*}
\tyrule{letimpl}
  { \Gamma \vdash e_1 : \ident{num} &
    \Gamma \vdash e_2 : \ctyp{\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \} }}{\tau} }
  { \Gamma \vdash \ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n \} }}~e_1~e_2 : 
      \ctyp{\cclrd{ \{ \ident{?p}_1, \ldots, \ident{?p}_n, \ident{?p} \} }}{\tau} }
\end{equation*}

\begin{equation*}
\semdef
  { \coctx{\Gamma}{\cclrd{ \{ \ident{?p} \} }} \vdash \ident{?p} : \ident{num} }
  {\lambda ctx.\ident{lookup}_{\cclrd{\ident{?p}}}~\ctx }
\end{equation*}

\begin{equation*}
\semdefff
  { \coctx{\Gamma}{\cclrd{r}} \vdash e_1 : \tau_1}
  { \coctx{\Gamma}{\cclrd{s}} \vdash e_2 : \tau_2}
  {\hspace{-1em}\begin{array}{l}  
      \coctx{\Gamma}{\cclrd{r} \;\cclrd{\cup\; (\cclrd{s} \setminus \{ \ident{?p}) \}} }\\[-0.3em]
        \quad \vdash \kvd{let}~\ident{?p}=e_1\\[-0.3em]
        \quyad\qquad \kvd{in}~e_2 : \tau_2 \\[-0.3em]
        ~ \\[-0.3em]
   \end{array}}
  {f}
  {g}
  {\hspace{-1em}\begin{array}{l}  
  \lambda\ctx.~~\\[-0.3em]
    \quad \kvd{let}~\ctx_0 = \ident{map}_{\cclrd{r} \;\cclrd{\cup}\; \cclrd{(s\setminus{\ident{?p}})}}~\ident{dup}~\ctx\\[-0.3em]
    \quad \kvd{let}~(\ctx_1, \ctx_2) = \ident{split}_{\cclrd{r}, \cclrd{(s\setminus{\ident{?p}})}}~\ctx_0\\[-0.3em]
    \quad g~(\ident{letimpl}_{\cclrd{\ident{?p}}, \cclrd{(s\setminus{\ident{?p}})}}~(f~\ctx_1)~\ctx_2)
  \end{array} }   
\end{equation*}

::: TODO ::: Everything except for counit

\begin{equation*}
\begin{array}{rl}
{\footnotesize(\emph{counit})} & \begin{array}{l}
\ident{counit}_{\cclrd{ \emptyset }}~\ident{Impl} (v, \langle \ldots \rangle)   \rightarrow v
\end{array}
\\[0.5em]
{\footnotesize(\emph{cobind})} & \begin{array}{l}
\ident{cobind}_{\cclrd{m}, \cclrd{n}}~f~(\ident{Df}\langle v_0, \ldots v_{m+n} \rangle) \rightarrow
\\[-0.25em]
\quad (\ident{Df}\langle f (\ident{Df}\langle v_0, \ldots, v_m \rangle), \ldots, f (\ident{Df}\langle v_{n}, \ldots, v_{m+n} \rangle) \rangle)
\end{array}
\\[0.5em]
{\footnotesize(\emph{merge})} & \begin{array}{l}
\ident{merge}_{\cclrd{m}, \cclrd{n}} ((\ident{Df}\langle v_0, \ldots, v_m \rangle), (\ident{Df}\langle v'_0, \ldots, v'_n\rangle)) \rightarrow
\\[-0.25em]
\quad (\ident{Df}\langle (v_0, b_0), \ldots, (v_{\mathit{min}(m,n)}, v'_{\mathit{min}(m,n)}) \rangle)
\end{array}
\\[0.5em]
{\footnotesize(\emph{split})} & \begin{array}{l}
\ident{split}_{\cclrd{m}, \cclrd{n}} (\ident{Df}\langle (v_0, b_0), \ldots, (v_{\mathit{max}(m,n)}, b_{\mathit{max}(m,n)}) \rangle)  \rightarrow
\\[-0.25em]
\quad \ident{Df}\langle v_0, \ldots, v_m \rangle, (\ident{Df}\langle b_0, \ldots, b_n\rangle
\end{array}
\\[0.5em]
{\footnotesize(\emph{prev})} & \begin{array}{l}
\ident{prev}_{\cclrd{n}} (\ident{Df}\langle v_0, \ldots, v_n, v_{n+1} \rangle) \rightarrow
\\[-0.25em]
\quad \ident{Df}\langle v_0, \ldots, v_n \rangle
\end{array}
\end{array}
\end{equation*}

\caption{Additional constructs embedding implicit parameters into the language}
\label{fig:semantics-ext-df}
\end{figure}
\newpage
